---
title: "COVAILworkflow"
output:
  rmarkdown::html_document:
  # rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    toc_float: true
vignette: >
  %\VignetteIndexEntry{COVAILworkflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## COVAIL Sample Manifest

The full COVAIL sample manifest is exported (.xlsx) from BLIS:
[COVAIL manifest (BLIS)](https://blis.urmc.rochester.edu/labkey/specimen/Topham%20Lab/COVAIL_Vaccine_Study/specimens.view?showVials=true#).

The manifest is pre-sorted on the column 'Global Unique Id' (ascending).

To ensure reproducibility, check the MD5 hash of the exported file; the historic value is:

**b00113dec261d816acd21aeddba8d183**

```{r sample.manifest.xlsx.hash}
manifest.BLIS.path<-list.files(
  system.file("extdata",package="COVAILworkflow"),
  pattern="SpecimenDetail", full.names = TRUE
  )
as.vector(tools::md5sum(manifest.BLIS.path)=="b00113dec261d816acd21aeddba8d183")

if(tools::md5sum(manifest.BLIS.path)!="b00113dec261d816acd21aeddba8d183"){
  stop("BLIS sample manifest MD5 hash does not match historic value.")
}
```

The manifest is modified and converted to a [data.table](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html).

* column names are made syntactically-valid
* ordered by *subject.id* and *visit*
* *subject.id* is re-encoded using [digests](https://cran.r-project.org/web/packages/digest/index.html)
* *visit* is re-encoded by dropping the conserved values (*323232##48*):
    + *3232324948* = 'V49'
    + *3232325148* = 'V51'
    + *3232325348* = 'V53'

```{r sample.manifest.xlsx.to.dt}
manifest<-COVAILworkflow::BLIS.manifest.prepare(manifest.BLIS.path)
manifest[sample.id %in% unique(sample.id)[1:3]]
```

## COVAIL Shipping Manifest

The modified COVAIL sample manifest is merged with the COVAIL shipping manifest to provide an exact physical location for each sample vial.

Check the MD5 hash of the file; the historic value is:

**9321b1e2920f72b8f511ffe74bc45834**

```{r shipping.manifest.xlsx.hash}
manifest.physical.path<-list.files(
  system.file("extdata",package="COVAILworkflow"),
  pattern="manifest_physical", full.names = TRUE
  )
as.vector(tools::md5sum(manifest.physical.path)=="9321b1e2920f72b8f511ffe74bc45834")

if(tools::md5sum(manifest.physical.path)!="9321b1e2920f72b8f511ffe74bc45834"){
  stop("Shipping manifest MD5 hash does not match historic value.")
}
```

* ordered by *subject.id* and *visit* and *box.row.col*

```{r sample.shipping.manifest.dt}
manifest<-COVAILworkflow::BLIS.manifest.prepare(
  manifest.BLIS.path,
  manifest.physical.path
)
manifest[sample.id %in% unique(sample.id)[1:3]]
```

## COVAIL Batch Assignment: Manifest

The merged COVAIL sample/shipping manifest is used to assign samples to pre-organized batches (randomized subject selection). Both a long-form and collapsed data.table are returned. The 'collapsed' form is used to pull batch-specific samples (vials) into a pre-mapped 'staging box' according to their pre-assigned 'batch.order'.  Once assigned, the manifest is saved (.xlsx) for future reference.

```{r sample.shipping.manifest.dt.batched,fig.width = 8,fig.height=8}
manifest.batched<-COVAILworkflow::manifest.batch.assign(
  manifest,
  vials.per = 3,
  add.controls = TRUE,
  workbook.path = NULL#set path here for saving .xlsx
)

##representative example: "COVAIL_001" pull/stage list
manifest.batched$manifest_collapsed[batch.seq=="001"]
stage.boxes<-COVAILworkflow::plot_manifest.stage(manifest.batched$manifest_collapsed)
##representative example: "COVAIL_001" stage box
stage.boxes[[1]]
```

## COVAIL Batch Assignment: Batch-Specific Sheets

A batch-assigned manifest is used to create batch-specific workbooks, each initially containing three sheets: a pull/stage list, a stage box plot (see above), and a 'Day 1: Thaw' sheet.  These sheets are designed to be printable to allow for efficient pulling and organization of individual batches.  In practice, a workbook is created for every batch, and then opened/printed using Excel -- here, a single representative batch ("COVAIL_001") is saved/shown.

```{r manifest.batched}
##a batched manifest was previously created
manifest.batched.path<-list.files(
  system.file("extdata",package="COVAILworkflow"),
  pattern="COVAIL_manifest_batched", full.names = TRUE
)
as.vector(tools::md5sum(manifest.batched.path)=="71a2f44bf9ef106720b4781ef2b9a9cd")

if(tools::md5sum(manifest.batched.path)!="71a2f44bf9ef106720b4781ef2b9a9cd"){
  stop("Batched manifest MD5 hash does not match historic value.")
}

##representative example: "COVAIL_001"; create a workbook
COVAILworkflow::COVAILworkbook.initialize(
  covail.manifest.batched.xlsx = manifest.batched.path,
  batch.name = "COVAIL_001",
  workbook.dir = tempdir(),
  .overwrite = TRUE
)

##representative example: "COVAIL_001"; a pull/stage list
data.table::setDT(
  openxlsx::read.xlsx(
    file.path(tempdir(),"COVAIL_001_workflow.xlsx")
  )
)[]

##representative example: "COVAIL_001"; Day 1: Thaw sheet
data.table::setDT(
  openxlsx::read.xlsx(
    file.path(tempdir(),"COVAIL_001_workflow.xlsx"),
    sheet = 3
  )
)[]
```

## COVAIL Day 1: Thaw

Consulting both the batch-specific pull/stage list and staging box, individual sample vials are 'pulled' and pre-organized; at the start of the batch-specific assay (Day 1: Thaw), samples are thawed into pre-labelled 15 mL tubes according to their pre-assigned batch order (see reference worksheet above; 'tube.label').

```{r COVAIL_001_tubes,echo=FALSE, fig.cap="Tube labels",out.width='50%',out.height='50%'}
tubes.path<-list.files(
  system.file("extdata",package="COVAILworkflow"),
  pattern="tubes.png", full.names = TRUE
)
suppressWarnings(knitr::include_graphics(tubes.path))
```

## COVAIL Day 2: Cell Counts and Stimulation

After overnight resting, cells are counted and the results (% viable, total viable cells) are used to determine per well stimulation plate mappings.  The cell counts are appended to existing metadata.

