---
title: "COVAIL Protocol:  \nStaining for Full Spectrum Cytometry  \nPanel: CYTOKINE\n"
author: "Nathan Laniewski"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
subtitle: 'Version: `r substr(tools::md5sum("./inst/protocols/COVAIL_protocol_staining.Rmd"),1,4)`'
always_allow_html: true
---

```{=html}
<style>
@media print{
  a[href]:after{
    content:""!important;
  }
}
</style>
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Reagents

```{r reagents.table}
COVAILworkflow:::reagents.table('kable')
```

#
1. Add 200 $\mu$L **SBn** to fresh V-bottom plate(s), matching the layout/wells of respective stim. plate(s)
2. Place V-bottom plate(s) and stim. plate(s) in a 37^o^C, 5% CO~2~ incubator for 10 minutes.
3. Remove **SBn** from V-bottom plate(s) and transfer cells from respective stim. plate(s)
    + NOTE: Pulse-pipette wells of stim. plate(s) to fully resuspend cells before transferring
4. Spin 300xg, 5 minutes, room-temperature
5. Multichannel aspirate supernatant
    + NOTE: Visually confirm that cell pellets are concentrated/compacted
6. Stain: **Barcodes (CD45)**; 50 $\mu$L, 20 minutes, room-temperature
    + NOTE: Pulse-pipette to mix; minimize the generation of air bubbles
7. *WASH*: 150 $\mu$L **SBn**; spin 300xg, 5 minutes, room-temperature; multichannel aspirate supernatant
8. *WASH*: 200 $\mu$L **SBn**; spin 300xg, 5 minutes, room-temperature; multichannel aspirate supernatant
9. *WASH*: 200 $\mu$L **SBn**; spin 300xg, 5 minutes, room-temperature; multichannel aspirate supernatant
10. Pool each **Barcode Block** into respective, pre-labeled 15 mL tube
    + NAMING CONVENTION: *BCpool_BLOCKnumber_PANEL*; e.g. *BCpool_BLOCK1_CYTOKINE*
11. Spin 300xg, 5 minutes, room-temperature
12. Vacuum aspirate supernatant
    + NOTE: leave ~20 $\mu$L residual
13. Stain: **Scarlet Viability Dye Mix**; 50 $\mu$L
    + NOTE: Fully disrupt cell pellet in remaining residual before staining
    + NOTE: Immediately before use, prepare **Scarlet Viability Dye Mix**; 1 $\mu$L dye + 3.2 mL **VB**
14. Spike-in: **TruStain FCx**; 5 $\mu$L
15. *Vortex-Genie Shaker*: 10 minutes, room-temperature
16. Spike-in: **Surface Mix**; 50 $\mu$L
    + NOTE: Pre-make surface mix
    + NOTE: Immediately before use, spin-filter **Surface Mix**; 12,000xg, 3 minutes
17. *Vortex-Genie Shaker*: 20 minutes, room-temperature
18. *WASH*: 2 mL **SBn**; spin 300xg, 5 minutes, room-temperature; vacuum aspirate supernatant
19. *WASH*: 2 mL **SBn**; spin 300xg, 5 minutes, room-temperature; vacuum aspirate supernatant
20. For each **Barcode Block/Pool**: pre-aliquot 500 $\mu$L of **BD Cytofix/Cytoperm** into a fresh 1.5 mL tube
21. For each **Barcode Block/Pool**: 
    1. Fully resuspend (ensure complete disruption) cell pellet in 50 $\mu$L **SBn**; pulse pipette a minimum of 15 times
    2. Without changing tip, rinse tip in respective, previously aliquoted **BD Cytofix/Cytoperm** to maximally recover residual cells
22. Spike-in: **BD Cytofix/Cytoperm** (respective); 500 $\mu$L
23. Immediately vortex (setting 5) for a full 10 second count
    + NOTE: ensures maximal, fixed single-cells
24. *Vortex-Genie Shaker*: 20 minutes, room-temperature
    + NOTE: prepare **ICS Mix**

  **!!!All subsequent spins: 800xg!!!**
    
25. *WASH*: 2 mL **PB**; spin 800xg, 5 minutes, room-temperate; vacuum aspirate supernatant
26. *WASH*: 2 mL **PB**; spin 800xg, 5 minutes, room-temperate; vacuum aspirate supernatant
27. Stain: **ICS Mix**; 50 $\mu$L
    + NOTE: Fully disrupt cell pellet in remaining residual before staining
28. *Vortex-Genie Shaker*: 30 minutes, room-temperature
29. *WASH*: 2 mL **PB**; spin 800xg, 5 minutes, room-temperate; vacuum aspirate supernatant
30. *WASH*: 2 mL **PB**; spin 800xg, 5 minutes, room-temperate; vacuum aspirate supernatant
    + NOTE: Before spinning, take an aliquot for counting: 5 $\mu$L of cells/pool + 15 $\mu$L **SB**
    + NOTE: Cellometer K2; assay: *1.6% fixed*; dilution factor: 4
31. Resuspend in **SB** at a concentration of 15E6/mL; store at 4^o^C (fridge)
    + NOTE: total cells/15E6 = resuspension volume (mL); e.g., 26.3E6 (total cells)/15E6 = 1.75 mL
