---
title: "COVAIL Protocol:  \nStaining for Full Spectrum Cytometry  \nPanel: CYTOKINE\n"
author: "Nathan Laniewski"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
always_allow_html: true
params:
  number.of.blocks.pools:
    value: x
---

```{=html}
<style>
@media print{
  a[href]:after{
    content:""!important;
  }
}
</style>
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r protocol.version}
##hash check body (.rmd file itself)
##hash check panel
protocol.hash.body<-tools::md5sum(
  list.files(
    system.file("protocols",package="COVAILworkflow"),
    pattern="COVAIL_protocol_staining_CYTOKINE.Rmd",
    full.names = TRUE
  )
)
protocol.hash.panel<-tools::md5sum(
  list.files(
    system.file("panels",package = "COVAILworkflow"),
    full.names = TRUE,
    pattern = "panels_COVAIL.csv"
  )
)
hashes<-lapply(list(body=protocol.hash.body,panel=protocol.hash.panel),function(i){
  paste(substr(i,1,4),substr(i,nchar(i)-3,nchar(i)),sep="...")
})
```
### Version:<br/>Body:   `r hashes$body`<br/>Panel:  `r hashes$panel`

# Reagents

```{r reagents.table}
COVAILworkflow:::reagents.table(reagent.type = 'CYTOKINE',table.type = 'kable')
```

# Buffers

```{r buffers.table}
COVAILworkflow:::buffers.table(buffers.type = 'CYTOKINE')
```

# Panels

## Panel: CD45 Barcodes
```{r barcodes.panel}
"Placeholder CD45 Barcodes"
# knitr::kable(list(COVAILworkflow:::panel.table()))
```

## Panel: Extracellular/Surface
Tests = 2.2
```{r extracellular.panel}
COVAILworkflow:::panel.kable()$buffer$CYTOKINE.extracellular
COVAILworkflow:::panel.kable()$mix$CYTOKINE.extracellular
```

## Panel: Intracellular
Tests = 2.2
```{r intracellular.panel}
COVAILworkflow:::panel.kable()$buffer$CYTOKINE.intracellular
COVAILworkflow:::panel.kable()$mix$CYTOKINE.intracellular
```

# Protocol

#
#. $\square$ Add 200 $\mu$L **SBn** to fresh V-bottom "stain" plate(s), matching the layout/wells of respective stim. plate(s)
#. $\square$ Place stain plate(s) and stim. plate(s) in a 37^o^C, 5% CO~2~ incubator for 10 minutes.
#. $\square$ Remove **SBn** from stain plate(s) and transfer cells from respective stim. plate(s)
    + NOTE: Pulse-pipette wells of stim. plate(s) to fully resuspend cells before transferring
#. $\square$ *WASH* stim. plate(s): 200 $\mu$L **SBn**; spin 300xg, 5 minutes, room-temperature
#. $\square$ Use 200 $\mu$L **SBn** from stim. plate(s) to *WASH* respective stain plate(s)
    + NOTE: Stain plate(s) -- visually confirm that cell pellets are concentrated/compacted
    + NOTE: Stim plate(s) -- resuspend any residual cells before adding to respective stain plate(s)
#. $\square$ Spin 300xg, 5 minutes, room-temperature; multichannel aspirate supernatant
#. $\square$ Stain: **Barcodes (CD45)**; 50 $\mu$L, 20 minutes, room-temperature
    + NOTE: Pulse-pipette to mix; minimize the generation of air bubbles
#. $\square$ *WASH*: 150 $\mu$L **SBn**; spin 300xg, 5 minutes, room-temperature; multichannel aspirate supernatant
#. $\square$ *WASH*: 200 $\mu$L **SBn**; spin 300xg, 5 minutes, room-temperature; multichannel aspirate supernatant
#. $\square$ *WASH*: 200 $\mu$L **SBn**; spin 300xg, 5 minutes, room-temperature; multichannel aspirate supernatant
#. $\square$ Pool each **Barcode Block** into respective, pre-labeled 15 mL tube
    + NAMING CONVENTION: *Study_Batch_BCpool_BLOCKnumber_PANEL*; 
    + Example: *COVAIL_001_BCpool_BLOCK1_CYTOKINE*
    + Example: *COVAIL_001_BCpool_BLOCK2_CYTOKINE*
#. $\square$ Spin 300xg, 5 minutes, room-temperature
#. $\square$ Vacuum aspirate supernatant
    + NOTE: leave ~20 $\mu$L residual
#. $\square$ Stain: **Scarlet Viability Dye Mix**; 50 $\mu$L
    + NOTE: Fully disrupt cell pellet in remaining residual before staining
    + NOTE: Immediately before use, prepare **Scarlet Viability Dye Mix**; 1 $\mu$L dye + 3.2 mL **VB**
#. $\square$ Spike-in: **TruStain FCx**; 5 $\mu$L
#. $\square$ *Vortex-Genie Shaker*: 10 minutes, room-temperature
    + NOTE: Spin-filter pre-made **Surface Mix**; 12,000xg, 3 minutes
#. $\square$ Spike-in: **Surface Mix**; 50 $\mu$L
#. $\square$ *Vortex-Genie Shaker*: 20 minutes, room-temperature
#. $\square$ *WASH*: 2 mL **SBn**; spin 300xg, 5 minutes, room-temperature; vacuum aspirate supernatant
#. $\square$ *WASH*: 2 mL **SBn**; spin 300xg, 5 minutes, room-temperature; vacuum aspirate supernatant
#. $\square$ For each **Barcode Block/Pool**: pre-aliquot 500 $\mu$L of **BD Cytofix/Cytoperm** into a fresh 1.5 mL tube
#. $\square$ For each **Barcode Block/Pool**: 
    1. Fully resuspend (ensure complete disruption) cell pellet in 50 $\mu$L **SBn**; pulse pipette a minimum of 15 times
    2. Without changing tip, rinse tip in respective, previously aliquoted **BD Cytofix/Cytoperm** to maximally recover residual cells
#. $\square$ Spike-in: **BD Cytofix/Cytoperm** (respective); 500 $\mu$L
#. $\square$ Immediately vortex (setting 5) for a full 10 second count
    + NOTE: ensures maximal, fixed single-cells
#. $\square$ *Vortex-Genie Shaker*: 20 minutes, room-temperature
    + NOTE: prepare **ICS Mix**
<br><br>**!!!All subsequent spins: 800xg!!!**<br><br>
#. $\square$ *WASH*: 2 mL **PB**; spin 800xg, 5 minutes, room-temperate; vacuum aspirate supernatant
#. $\square$ *WASH*: 2 mL **PB**; spin 800xg, 5 minutes, room-temperate; vacuum aspirate supernatant
#. $\square$ Stain: **ICS Mix**; 50 $\mu$L
    + NOTE: Fully disrupt cell pellet in remaining residual before staining
#. $\square$ *Vortex-Genie Shaker*: 30 minutes, room-temperature
#. $\square$ *WASH*: 2 mL **PB**; spin 800xg, 5 minutes, room-temperate; vacuum aspirate supernatant
#. $\square$ *WASH*: 2 mL **PB**; spin 800xg, 5 minutes, room-temperate; vacuum aspirate supernatant
    + NOTE: Before spinning, take an aliquot for counting: 5 $\mu$L of cells/pool + 15 $\mu$L **SB**
    + NOTE: Cellometer K2; assay: *1.6% fixed*; dilution factor: 4
#. $\square$ Resuspend in **SB** at a concentration of 15E6/mL; store at 4^o^C (fridge)
    + NOTE: total cells/15E6 = resuspension volume (mL); e.g., 26.3E6 (total cells)/15E6 = 1.75 mL
